Sequential rank agreement methods for comparison of ranked lists

### 2016-06-17 latex/superranker-manuscript.tex
Fixed part II

We now define the parameter of interest. For any given deterministic threshold function $q\colon\, 1,\ldots,P \mapsto \mathbb{R}_{\geq 0}$
we define the population superlevel set of the sequential rank agreement as
\begin{align}
  \mathcal{L}(q) = \left\{d : \textrm{sra}(d) \geq q(d)\right\}.
\end{align}
The values of $q$ could for example be the limits-of-agreement for randomly
permuted lists as in the null-hypothesis in equation (\ref{eq:permutationHypothesis}).
A change-point $d^\ast(q)$ in the list agreement is then defined by the position
\begin{align}
  d^\ast(q) &= 1\left(|\mathcal{L}(q)| > 0\right)\inf \mathcal{L}(q) + 1\left(|\mathcal{L}(q)| = 0\right)P
\end{align}
corresponding to the first list depth where the sequential rank agreement exceeds the threshold 
if such a position exists. Otherwise, the full list is in agreement according to $q$ and the change-point
is defined as the length of the lists. The empirical superlevel set is similarly defined as
\begin{align}
 \widehat{\mathcal{L}}_L(\widehat{q}_L) = \left\{d : \widehat{\textrm{sra}}_L(d) \geq \widehat{q}_L(d)\right\}
\end{align}
where we allow the threshold function to depend on the sample size as well. The estimated change-point is therefore
\begin{align}
  \widehat{d^\ast_L}(\widehat{q}_L) &= 1(|\widehat{\mathcal{L}}_L(\widehat{q}_L)| > 0)\inf \widehat{\mathcal{L}}_L(\widehat{q}_L) + 1(|\widehat{\mathcal{L}}_L(\widehat{q}_L)| = 0)P
\end{align}

The consistency of the estimated change-point, $\widehat{d^\ast_L}(\widehat{q}_L)$, follows from Theorem \ref{thm:consistency} by the following corollary. 
\begin{corollary}
Let $\widehat{q}_L$ be a threshold function such that $\left\|\widehat{q}_L - q\right\|_\infty = o_P(1)$ for some limiting
function $q$.  Then, $\widehat{d^\ast_L}(\widehat{q}_L) \overset{P}{\longrightarrow} d^\ast(q)$ for $L \rightarrow \infty$.
\label{cor:consistency}
\end{corollary}

\mcomment{Elaborate more on how to apply this in practice (finite
  sample/bootstrap/etc.) and that $q$ is similar to LoA BA-style? Show
some result althrough }

@@

In conclusion we have introduced a method for evaluation of ranked
(censored) lists that can be easily interpreted and that can be
applied to a large number of situations.  The method presented here
can be adapted further by using it to compare and classify statistical
analysis methods that agree on the rankings they provide or by using
the rank agreement to optimize a hyper-parameter in, say, elastic net
regularized regression where the rank agreement is used to determine
the mixing proportion between the $L_1$ and the $L_2$ penalty.
%Finally, the proposed method may be adapted (with some additional
%assumptions) to the situation where the lists are subsets of each
%other.
We will be investigating these extensions further in the future.

>>>>

We now define the parameter of interest. For any given deterministic threshold function $q\colon\, 1,\ldots,P \mapsto \mathbb{R}_{\geq 0}$
we define the population superlevel set of the sequential rank agreement as
\begin{align}
  \mathcal{L}(q) = \left\{d : \textrm{sra}(d) \geq q(d)\right\}.
\end{align}
The values of $q$ could for example be the limits-of-agreement for randomly
permuted lists as in the null-hypothesis in equation (\ref{eq:permutationHypothesis}).
A change-point $d^\ast(q)$ in the list agreement is then defined by the position
\begin{align}
  d^\ast(q) &= 1\left(|\mathcal{L}(q)| > 0\right)\inf \mathcal{L}(q) + 1\left(|\mathcal{L}(q)| = 0\right)P
\end{align}
corresponding to the first list depth where the sequential rank agreement exceeds the threshold 
if such a position exists. Otherwise, the full list is in agreement according to $q$ and the change-point
is defined as the length of the lists. The empirical superlevel set is similarly defined as
\begin{align}
 \widehat{\mathcal{L}}_L(\widehat{q}_L) = \left\{d : \widehat{\textrm{sra}}_L(d) \geq \widehat{q}_L(d)\right\}
\end{align}
where we allow the threshold function to depend on the sample size as well. The estimated change-point is therefore
\begin{align}
  \widehat{d^\ast_L}(\widehat{q}_L) &= 1(|\widehat{\mathcal{L}}_L(\widehat{q}_L)| > 0)\inf \widehat{\mathcal{L}}_L(\widehat{q}_L) + 1(|\widehat{\mathcal{L}}_L(\widehat{q}_L)| = 0)P.
\end{align}

The consistency of the estimated change-point, $\widehat{d^\ast_L}(\widehat{q}_L)$, follows from Theorem \ref{thm:consistency} by the following corollary. 
\begin{corollary}\label{thecor}
Let $\widehat{q}_L$ be a threshold function such that $\left\|\widehat{q}_L - q\right\|_\infty = o_P(1)$ for some limiting
function $q$.  Then, $\widehat{d^\ast_L}(\widehat{q}_L) \overset{P}{\longrightarrow} d^\ast(q)$ for $L \rightarrow \infty$.
\label{cor:consistency}
\end{corollary}

Corollary \ref{thecor} tells us that we can use the a threshold
function $\widehat{q}_L$ estimated under the null hypothesis as
discussed in the previous section as a limiting threshold function for
inferring the depth $d$, where the observed sequential rank agreement
first crosses the threshold of the null threshold, \ie, the depth
until which the observed ranked lists are in better agreement than
expected under the null hypothesis. In that sense the threshold
function serves the same role as the limits of agreement in method
comparison studies, except that the threshold funciton is not constant
but can accommodate the changing nature of the number of items used
for the computation of the equential rank agreement for a given depth.

In practice we can compute an estimate of the threshold function under
the null using the permutation approach sketched in the previous
section which makes it relevant even in small sample settings.

\mcomment{Elaborate more on how to apply this in practice (finite
  sample/bootstrap/etc.) and that $q$ is similar to LoA BA-style?  }

@@

In conclusion we have introduced a method for evaluation of ranked
(partial/censored) lists that can be easily interpreted and that can be
applied to a large number of situations.  The method presented here
can be adapted further by using it to compare and classify statistical
analysis methods that agree on the rankings they provide or by using
the rank agreement to optimize a hyper-parameter in, say, elastic net
regularized regression where the rank agreement is used to determine
the mixing proportion between the $L_1$ and the $L_2$ penalty.
%Finally, the proposed method may be adapted (with some additional
%assumptions) to the situation where the lists are subsets of each
%other.
We will be investigating these extensions further in the future.
